## Introduction to Numeric Vectors {#intro-num-vector} 

We will start off this chapter by learning **numeric vectors**. Numeric vectors are perhaps the most commonly used member of the **atomic vector** family, where all elements are of the same type. 

### Numeric vectors: creation and class{#create-numeric-vector} 

```{r, include = FALSE, echo=FALSE}
rm(list=ls())
```

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>", echo = TRUE, eval = TRUE, results = FALSE, error = TRUE, fig.width = 6, warning = FALSE, collapse = TRUE)
```

A **numeric vector** is an atomic vector containing only numbers. For example, `6` is a numeric vector with one element of value 6. 
 

By assigning the value 6 to the name x1, you have created a new numeric vector `x1`. As a result, you can refer to `x1` in subsequent calculations. For any vector, you can use the `length()` function to check the number of elements it contains. Given the output, you see that have successfully created the numeric vector `x1` of length 1.
 
```{r, results = TRUE}
6                         #a numeric vector with length 1
x1 <- 6                   #x1 is also a numeric vector
x1                        #check the value of x1
length(x1)                #length of x1
```

Moving on, you may wonder, can a numeric vector contain more than one values? The answer is a big YES! In R, you can use the `c()` function (`c` is short for combine) to combine elements into a numeric vector. 

```{r, results = TRUE}
c(1, 3, 3, 5, 5)          #use c() to combine elements into a numeric vector of length 5
y1 <- c(1, 3, 3, 5, 5)    #y1 is a numeric vector of length 5
y1                        #check the value of y1
length(y1)                #length of y1
```

In this example, you have created a length-5 object using the `c()` function with arguments being the five elements separated by comma. Since the value of each element is a number, the object is a numeric vector. 

Notice that the second and third elements in `y1` have the same value 3. Similar to `x1`, you can verify the contents of `y1` and check the length of it via the `length()` function.

::: {.infobox .caution data-latex="{caution}"}

When you assign several values to a name, the order of the values will not change after assignment. If you create two numeric vectors containing the same numbers but in different orders, the two vectors will maintain the specified orders. For example,

```{r, results = TRUE}
y2 <- c(1, 3, 5, 7, 9)    
y2                        
y3 <- c(9, 7, 5, 3, 1)    
y3
```

::: 

In addition to using numbers inside the `c()` function, you can also use several numeric vectors as the arguments to create a longer vector. The new. longer vector will combine the input numeric vectors in the given order. 

```{r, results = TRUE}
c(x1, y1)          #use c() to combine several numeric vectors into one numeric vector
z1 <- c(x1, y1)
z1
length(z1)
```



For any R object, you can use the function `class()` to check its **class**. A class can be thought of as a "type," providing a description about the object, and determines what functions can be applied to it. 

```{r, results = TRUE, collapse = TRUE}
class(x1)
class(y1)
class(z1)
```

From the results, you can see that `x1`, `y1`, and `z1` are all numeric, which is the reason why they are called *numeric vectors*. 


### Numeric vectors: access and modify elements

To **access an element** from a vector, you can use the vector name followed by a pair of `[` and `]`containing the index of the desired element. Let's see some examples.

```{r, results = TRUE}
y1[2]     #access the second element of y1
y2[3]     #access the third element of y2
```


You can also **modify** a particular element of a vector by using the *assignment operator* with the access expression on the left and the new value on the right.
```{r, results = TRUE}
y1               #the current value of y1
y1[2] <- 100     #modify the second element of y1 to 100
y1               #the new value of y1
```

As you can see here, the second element of `y1` is modified to 100, which is reflected via checking the value of `y1`. 

Let's look at another example where we want to modify the third element of `y2` to be twice as much as the fourth element of `y1`.
```{r, results = TRUE}
y2                     #the current value of y2
y2[3] <- y1[4] * 2     #modify the third element of y2
y2                     #the new value of y2
```




### Numeric vectors: operations and recycling rule

Since numeric vectors are purely made of numbers, you can do **arithmetic operations** between them, just like the fancy calculator in Section \@ref(Calculator). If two vectors are of the **same length**, the operation is done **element-wisely**. In other words, R will perform the operation between elements in the same index of different vectors. First, let's create another vector `x2` of length 1 and compute the sum of `x1` and `x2`. Also recall that we've previously assigned x1 to a length-1 numeric vector of value 6.

```{r results=TRUE}
x1
x2 <- 3
x1 + x2
```

Similarly, you can create another vector `y2` of the same length as vector `y1`. Then, you can do operations between `y1` and `y2`.

```{r results=TRUE}
y1
y2 <- c(2, 4, 1, 3, 2)
y1 * y2
```

The result is yet another length-5 vector. To check the calculation was indeed done element-wisely, you can verify that the value of the first element is $1 * 2 = 2$, and value of the second element is $100 * 4 = 400$, etc. 

Since the calculation is done element-wisely, we normally would want the two vectors to have the same length. However, there is an important **recycling** rule in R, which is quite useful and enables us to write simpler code. Specifically, if one vector is shorter than the other vector, R will **recycle** (repeat) the shorter vector until it matches in length with the longer one. This recycling is most often used for an operation between a **length>1** vector and a **length-1** vector. Let's see an example.

```{r results=TRUE}
y1 + x1
```

From the result, you can see that `x1` is recycled five times to match in length with `y1`. Consequently, each element in `y1` is added by 6.

The followings are a few additional examples you can try.

```{r, eval=FALSE}
y1 * x2
y1 / 5
y2 - x1
```


### Numeric vectors: storage types (doubles and intergers) {#storage-type}

Now, it is time to learn how numeric vectors are stored in R. To find the **internal storage type** of an R object, you can use the `typeof()` function.

```{r, results=TRUE}
my_double <- c(1, 3, 4)
typeof(my_double)         #internal storage type
```

We can see that the internal storage type of `my_num` is **double**, meaning that `my_num` is stored as a **double precision** numeric value. Looking at the values of `my_num`, it is easy to see that they are all integers. In fact, R stores numeric vectors as double precision vectors by default. That being said, you can always store the integers in a **integer type**, which offers great memory savings compared to doubles. The tricky part is that you usually need to explicitly tell R that you are storing them as integers. 

To create an integer vector, you can still use the `c()` function with the integers separated by comma as arguments. However, you need to put an "L" after each integer. Let's create an integer and check its storage type as well as its class. 
```{r, results=TRUE}
my_int <- c(1L, 3L, 4L)
typeof(my_int)
class(my_int)
```

You can see that `my_int` is indeed of `integer` type, with the `class` of it being `integer` as well. 

It is also worth noting that the displaying value of `my_double` and `my_int` are the same, though they are stored differently in the memory.
```{r, results=TRUE}
my_double
my_int
```


::: {.infobox .caution data-latex="{caution}"}
Despite the differences between integers and doubles, you can usually ignore their differences unless you are working on a very big data set. R will automatically convert objects between integers and doubles when necessary. 
:::

### Numeric vectors: printing

Now, you have already learned to run the object name to reveal its value. Let's try to type `pi`, which is an internal constant available for use.
```{r, results=TRUE}
pi
```

As you can see from the output, R prints out 7 significant digits by default, though in fact we need infinitely many digits to faithfully represent `pi`. To print out an object with a customized significant digit number, you can use the `print()` function that contains useful argument called `digits`, which controls the minimal number of significant digits. Let's see the following examples.
```{r, results = TRUE}
print(pi)
print(pi, digits = 3)            #print pi for 3 significant digits
print(pi, digits = 16)           #print pi for 16 significant digits
print(c(pi, exp(1), log(2)), digits = 4)
```

As you can imagine, the `print()` function will be very useful in creating tables that look more streamlined.


### Exercises

Write R code to complete the following tasks.

1. Create a numeric vector named `vec_1` with values $(2, 4, 6, 8)$, get its length,  find out its class, and get its storage type.

2. For the numeric vector `vec_2 <- c(1, 3, 7, 10)`, get the value of the 3rd element, multiple the 3rd element by 5, and verify the change. 

3. Create a vector `vec_3` where each element is twice the corresponding element  in `vec_1` minus half the corresponding element in `vec_2`. 

4. Create an integer vector `int_1` that contains integers $(2, 4, 6, 8)$. Check its class and storage type.

5. Print out the vector $(e, e^2, e^3)$ with 5 significant digits. 


