## Customization of Labels and Titles {#customization}

In Section \@ref(customization-axes), we discussed the customization of the x and y axes in ggplot. Now, we will introduce other types of customization, and the concept of themes.

Let's first review the scatterplot we created between `liv_area` and `sale_price` and save it into as an object `g`.

```{r, results=FALSE}
library(r02pro)
library(tidyverse)
g <- ggplot(data = sahp, mapping = aes(x = liv_area, y = sale_price)) + 
  geom_point()
```

***a. customize x and y labels and title***

By default, the x and y labels are the variables names, and there is no title for the plot. You can customize the x and y labels using the `xlab()` and `ylab()` function, and add a title with the `ggtitle()` function. 
```{r, warning=FALSE}
g1 <- g + 
  xlab("Living Area") + 
  ylab("Sale Price") + 
  ggtitle("Price vs. Area")
g1 
```

***b. customize the font of the x and y breaks***

In addition, you can further customize the font of the x and y breaks using the `theme()` function with the argument `axis.text`.  

```{r}
g1 + theme(axis.text = element_text(size = 25, color = "red"))
```


***c. customize the font of labels***

To customize the font, you can use `axis.title` argument to change the size, color, and face of the labels. 
```{r}
g1 + theme(axis.title = element_text(size = 18, color = "red", face = "italic"))
```



***d. customize the font of the title***

Similarly, you can use the `plot.title` argument to customize the font of the title. 

```{r}
g1 + theme(plot.title = element_text(size = 24, color = "magenta", face = "bold"))
```


***e. center the title***

Sometimes, we may want to center the title. We can achieve this by setting the `hjust` parameter. 

```{r}
g1 + theme(plot.title = element_text(size = 24, color = "magenta", face = "bold",hjust = 0.5)) 
```


***f. mix***

Apparently, you are free to mix all the different customizations. Let's see an example as below.
```{r}
g1 + theme(axis.title = element_text(size = 18, color = "red", face = "italic"), axis.text = element_text(size = 12, color = "blue"), plot.title = element_text(size = 24, color = "magenta", face = "bold", hjust = 0.5))
```

***g. save as a theme***

As you can see from the code, the code gets complicated if we want to customize many things at the same time. To save time, you can actually save the desired into an R object and reuse it later. 
```{r}
mytheme <- theme(axis.title = element_text(size = 18, color = "red", face = "italic"), axis.text = element_text(size = 12, color = "blue"), plot.title = element_text(size = 24, color = "magenta", face = "bold", hjust = 0.5))
```
Then, we can generate the same plot with `mytheme` using

```{r, eval=FALSE}
g1 + mytheme
```

Similarly, you can add another layer of smoothline fit and apply `mytheme`.

```{r, eval=FALSE}
g1 + 
  geom_smooth() + 
  mytheme
```

You may notice that title is not accurate. So, we want to change the title to reflect the addition of the smoothline fit. Let's simply add another `ggtitle()` function to overwrite the existing one in the theme.
```{r, eval=FALSE}
g1 + 
  geom_smooth() + 
  mytheme + 
  ggtitle("Scatterplot + smoothline") 
```



