## Density Plots {#density-plot}

In Section \@ref(histogram), we have learned to use `geom_histogram()` as a way to visualize the distribution of a continuous variable. In addition, we can also use it to generate a piece-wise constant estimate of the probability density function. Today, we will introduce another visualization method for continuous data, namely the **density plots**. First, let's review the `geom_histogram()` for estimating the density function. 

```{r, warning=FALSE}
library(ggplot2)
library(r02pro)
ggplot(data = sahp) +  
  geom_histogram(aes(x = sale_price, y = ..density..))
```

There are 30 bins by default. You may notice that this density estimate is not smooth, sometimes we may prefer a smoothed estimate. Then, we can use the `geom_density()` function to achieve this. 

```{r, warning=FALSE}
ggplot(data = sahp) + 
  geom_density(aes(x = sale_price))
```

This plot shows the so-called "kernel density estimate", a popular way to estimate the probability density function from sample. The density estimate can be viewed as a smoothed version of the histogram. We can combine the two plots together using global mapping. 

```{r, warning=FALSE}
ggplot(data = sahp, aes(x = sale_price)) +
  geom_histogram(aes(y = ..density..)) +
  geom_density(color = "red", size = 2)
```

Here, we added some global aesthetics in `geom_density()` to make the density plot red and the line width by setting `size = 2`. It is clear that the density plot is a useful alternative to the histogram for visualizing continuous data. 

### Aesthetics in Density Plots

Now, let's introduce some commonly used aesthetics for density plots.

***a. Color***

```{r, warning=FALSE}
ggplot(data = remove_missing(sahp, vars = "oa_qual")) +  
  geom_density(aes(x = sale_price, color = oa_qual > 5))
```

Here, we divide the data into two groups according to the value of `oa_qual`, then generate separate density estimates with different colors for `oa_qual > 5` and `oa_qual <= 5`. The blue curve represents the density estimates for larger values of `oa_qual`  while the red curve corresponds to that of the houses with smaller values.  

***b. Fill***

Another way to generate different density estimates is to use the `fill` aesthetic. Let's see the following example.

```{r, warning=FALSE}
ggplot(data = remove_missing(sahp, vars = "oa_qual")) +  
  geom_density(aes(x = sale_price, fill = oa_qual > 5))
```

The `fill` aesthetic also divides the data into groups according to `oa_qual`, then generate separate density estimates. The difference between  `fill` and  `color` aesthetics is that `fill`  generates shaded areas below each density curve with different colors while `color` generates density curves with different colors.  As we can see from the plot, there is a substantial overlap of the shaded areas. To fix this issue, we can change the transparency of the shades by adjusting the value of the `alpha` aesthetic.

```{r, warning=FALSE}
ggplot(data = remove_missing(sahp, vars = "oa_qual")) +  
  geom_density(aes(x = sale_price, fill = oa_qual > 5), 
               alpha = 0.5)
```

We can now see both shaded areas in a clearly way. 

***c. Linetype***

We can also use different linetypes for different curves.

```{r, warning=FALSE}
ggplot(data = remove_missing(sahp, vars = "oa_qual")) +  
  geom_density(aes(x = sale_price, linetype = oa_qual > 5))
```


***d. Global aesthetics***

As usual, we can also set global aesthetics for `geom_density()` and combine it with the mapped aesthetics. 

```{r, warning=FALSE}
ggplot(data = remove_missing(sahp, vars = "oa_qual")) +  
  geom_density(aes(x = sale_price, linetype = oa_qual > 5), 
               size = 1, 
               color = "red")
```

Here, the `size` controls the width of the density curve. 


### Exercises
Use the `sahp` data set to answer the following questions.

1. Create density plot on the living area (`liv_area`) with dashed lines and different colors for different values of `kit_qual`. What conclusions can you draw from the plot? 

2. Try to create density plot for `kit_qual`. Do you think this plot is informative? If not, create a plot that captures the distribution of `kit_qual`.   


