--- 
title: "R Programming: Zero to Pro"
author: "[Yang Feng](https://yangfeng.hosting.nyu.edu/) and [Jianan Zhu](https://yangfeng.hosting.nyu.edu/authors/jianan-zhu/)"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: "r02pro/r02pro.github.io"
colorlinks: yes
description: "R Programming: Zero to Pro"
---
```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>", echo = TRUE, eval = TRUE, results = FALSE, error = TRUE, fig.width = 6, warning = FALSE, collapse = TRUE
)
```

# Preface {-}

This book is for anyone who is interested in learning R and Data Science. It is designed for people with zero background in programming. 

We also have a companion R package named **r02pro**, containing the data sets used as well as interactive exercises for each part. 

# Contributing to the Book  {-}

To improve the book, we would like to ask for your help. 

### Submitting Questions or Feedback  {-}

If you have any questions or feedback (including typos or grammar issues) about any materials in the book, we greatly appreciate if you can write to us at r02pro007@gmail.com. We will also add your name in the Acknowledgment section to show our gratitude. 

### Contributing Exercises  {-}

We warmly welcome everyone to help the book via contributing exercises for each section. Please make sure the exercise you are contributing only needs the knowledge up to the specific section number (it of course can use any knowledge of all sections up to the particular section) you would like to put the exercise into.

Please submit the proposed exercise(s) via the  [Google Form](https://docs.google.com/forms/d/e/1FAIpQLSckpcPZvRSr2vju-J2w8NhOx_oIyYtd8uuX0xqq_ThfRWMltw/viewform).

Thanks a lot for your contribution.


# Acknowledgement {-}



```{r, results = "asis", echo = FALSE, message = FALSE}
library(dplyr)
contributors <- readr::read_csv("contributors.csv", col_types = list())

contributors <- contributors %>% 
  mutate(
    link = glue::glue("[\\@{login}](https://github.com/{login})"),
    desc = ifelse(is.na(login), name, ifelse(is.na(name), link, glue::glue("{name} ({link})")))
  )

cat("This book is the product of numerous collaborative efforts. Among all people, we would like to thank Dr. Rebecca Betensky, Dr. Rumi Chunara, and Dr. Hai Shu, for the initial planning of the new courses GPH-GU 2183: Introduction to Statistical Programming in R and GPH-GU 2183: Intermediate Statistical Programming in R. In addition, we
    are most grateful to the students in the course GPH-GU 2183: Introduction to Statistical Programming in R in Fall 2021 at New York University. The list of people that made contributions include   ")
cat(paste0(contributors$desc, collapse = ", "))
```

# Building Environment {-}

The R session information when compiling this book is shown below.
```{r, results=TRUE}
sessionInfo()
```


In the book, package names are in **bold** text (e.g., **r02pro**), and inline codes and functions are formatted in a typewriter font (e.g. `1 + 2`, `r02pro()`). We always add parentheses after a function name (e.g. `r02pro()`).
