## Global Aesthetics in ggplot {#global-aes}

Knowing how to generate a scatterplot using `geom_point()`, let's discuss one of the most important aspects in a `geom`, namely, the **aesthetics**. Aesthetics include various parameters that you can change that affect the appearances of a plot. Some commonly used aesthetics include color, size, shape, and so on. 

Note that although we will introduce aesthetics via the example of scatterplot, they are used for all kinds of plots which will be covered at a later time. 



In this section, we discuss **global aesthetics**, which change certain features of a plot globally, while in the next section, we discuss **aesthetics** mapping, which maps variables to certain aesthics. 


Let's first review the code we used to generate the scatterplot between `liv_area` and `sale_price`.

```{r}
library(ggplot2)
library(r02pro)
ggplot(data = sahp) + 
  geom_point(mapping = aes(x = liv_area, y = sale_price))
```


Now, let's see how to set **global aethetics** in `geom_point()`. 

***a. Color***

To change the color of all points, you can set the `color` argument in the `geom_point()` function. Note that it is placed outside of the `aes()` function.

```{r, results = TRUE}
ggplot(data = sahp) + 
  geom_point(mapping = aes(x = liv_area, y = sale_price), 
             color = "red")
```

Clearly, all points are changed to red. 


***b. Size***

Similarly, you can set the `size` element in the `geom_point()` function to change the size of the all points.

```{r, results = TRUE}
ggplot(data = sahp) + 
  geom_point(mapping = aes(x = liv_area, y = sale_price), 
             size = 3)
```

You may notice that the points are now bigger than before. Looking at the plot, many points are overlapping with each other, which is sometimes called **overplotting**. To solve this issue, you can change the transparency level of the points by setting the `alpha` argument. 

***c. Transparency***

```{r, results = TRUE}
ggplot(data = sahp) + 
  geom_point(mapping = aes(x = liv_area, y = sale_price), 
             size = 3, 
             alpha = 0.5)
```

By setting `alpha = 0.5`, the points become more visible and the overplotting problem is largely alleviated. 

***d. Shape***

Lastly, we can also change the shape of the points from the default one (circle) to other shapes by the `shape` argument in `geom_point()`. 

```{r, results = TRUE}
ggplot(data = sahp) + 
  geom_point(mapping = aes(x = liv_area, y = sale_price), 
             shape = 2)
```


***e. Multiple Aesthetics***

Of course, we can combine multiple global aesthetics in the same plot. 
```{r}
ggplot(data = sahp) + 
  geom_point(mapping = aes(x = liv_area, y = sale_price), 
             color = "red", 
             size = 3, 
             shape = 2)
```

Here, we have all points red, size of 3, and of triangle shape.



### Exercises

Using the `sahp` dataset with the **ggplot2** package, answer the following questions.


1.	Create a scatterplot to visualize the relationship between `lot_area` (on the x-axis) and `sale_price` (on the y-axis).
2.	In the scatterplot from Q1, change the size of all points to 3 and use different colors according to the value of `house_style`.
3.  Change legend order in the scatterplot from Q2 to be "1Story", "SFoyer", "1.5Fin", "2Story", "SLvl" from top to bottom.
4. In the scatterplot from Q1, change the shape of all points to triangle and map `bedroom` to the color aesthetic. Then map the low value of `bedroom` to the yellow color and map the high value of `bedroom` to the green color.
5. In the scatterplot from Q1, change the color of all points to green and use different shapes to distinguish whether the house has more than 3 bedrooms.




